---
title: "statlearning-final"
author: "Horatiu"
date: '2022-07-13'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

FIFA 22
first we do libraries

```{r}
library(knitr) 
#library(kableExtra) #nice html tables
#library(magick)     #working with images
library(ggplot2)    #graphs
library(reshape2)   #dataframe reshaping
library(viridis)    #colors
library(stringr)    #working with strings
library(FactoMineR) #factor analysis (PCA)
library(factoextra) #factor analysis (PCA)
#library(kohonen)    #self-organised maps/SOM 

# libraries
library(dplyr)
library(ggplot2)
library(MASS)
#library(randomForest)
library(class)
#library(caret)
install.packages("corrplot")

library(corrplot)
library(gridExtra)
remove.packages(rlang)
install.packages("rlang")
install.packages("https://cran.rstudio.com/bin/windows/contrib/4.1/caret_6.0-92.zip")
install.packages("caret")
#library(mlbench)
library(reshape)
library(corrplot)
library(caret)
library(randomForest)

set.seed(123)
```

## Data Cleaning
check dimensions, check columns

```{r}
players_full <- read.csv("E:/horatiu/FIS/players_22.csv/players_22.csv") #full dataframe
dim(players_full) #full dataset
colnames(players_full)
#summary(players_full)
```
###select relevant columns 
we also remove position stats because we want to predict positions

```{r}
players_22 <- subset(players_full, select = c("short_name","player_positions","age","height_cm","weight_kg","pace","shooting","passing","dribbling","defending","physic","attacking_crossing","attacking_finishing","attacking_heading_accuracy","attacking_short_passing","attacking_volleys","skill_dribbling","skill_curve","skill_fk_accuracy","skill_long_passing","skill_ball_control","movement_acceleration","movement_sprint_speed","movement_agility","movement_reactions","movement_balance","power_shot_power","power_jumping","power_stamina","power_strength","power_long_shots","mentality_aggression","mentality_interceptions","mentality_positioning","mentality_vision","mentality_penalties","mentality_composure","defending_marking_awareness","defending_standing_tackle","defending_sliding_tackle"))
dim(players_22)
# subset 38:72

```
```{r}
head(players_22, n=5)
summary(players_22)
```
## Check how many goalkeepers

```{r}
install.packages("stringr")     # Install & load stringr package
library("stringr")
goalkeepers <- str_detect(players_22$club_position, "GK")
sum(goalkeepers)

```
#eliminate goalkeepers
```{r}
players_22<-subset(players_22, player_pos!="GK")
which(apply(X = players_22, MARGIN = 2, FUN = anyNA) == TRUE) # check for NA
dim(players_22)
#head(players_22, n = 2)
```

```{r}
players_22 <- na.omit(players_22) # delete NA
dim(players_22)
col<-colnames(players_22)
col
```
### Now we take only the main position of each player

```{r}
players_22$player_positions<- word(players_22$player_positions, 1, sep = fixed(","))
unique(players_22$player_positions)
#insert picture of player positions
# now we group them into the main 9 positions

#central back players
players_22$player_positions[players_22[,"player_positions"]== "LCB"|players_22[,"player_positions"]== "CB"|players_22[,"player_positions"]== "RCB"] <- "CB"

#left back players
players_22$player_positions[players_22[,"player_positions"]== "LWB"|players_22[,"player_positions"]== "LB"]<-"LB"

#right back players
players_22$player_positions[players_22[,"player_positions"]== "RWB"|players_22[,"player_positions"]== "RB"]<-"RB"

#central defender midfielders
players_22$player_positions[players_22[,"player_positions"]== "LDM"|players_22[,"player_positions"]== "CDM"|players_22[,"player_positions"]== "RDM"] <- "CDM"

#central midfielders
players_22$player_positions[players_22[,"player_positions"]== "LCM"|players_22[,"player_positions"]== "CM"|players_22[,"player_positions"]== "RCM"] <- "CM"

#central attacking midfielder
players_22$player_positions[players_22[,"player_positions"]== "LAM"|players_22[,"player_positions"]== "CAM"|players_22[,"player_positions"]== "RAM"] <- "CAM"

#left wingers
players_22$player_positions[players_22[,"player_positions"]== "LM"|players_22[,"player_positions"]== "LW"|players_22[,"player_positions"]== "LF"] <- "LW"

#right wingers
players_22$player_positions[players_22[,"player_positions"]== "RM"|players_22[,"player_positions"]== "RW"|players_22[,"player_positions"]== "RF"] <- "RW"

#strikers
players_22$player_positions[players_22[,"player_positions"]== "LS"|players_22[,"player_positions"]== "CF"|players_22[,"player_positions"]== "RS"] <- "ST"

colnames(players_22)
```
#normalize numerical values
```{r}
# normalization function
normalize <-function(x) { (x -min(x))/(max(x)-min(x))   }

# normalize 
players_norm <- as.data.frame(lapply(players_22[, c(3:40)], normalize))

```

#check correlation matrix
```{r}
#create numerical subset
cormatrix <- cor(players_norm)
corrplot(cor(players_norm), method = 'shade', type = 'lower', order = 'hclust' )
```
```{r}
highcorr <- findCorrelation(cormatrix, cutoff=0.85)
highcorr
col2<-colnames(players_norm)
col2
col2<-col2[-highcorr]
corrplot.mixed(cor(players_norm[highcorr]), lower = "number", upper="pie") # it looks much better so we use the variables from col 2
```
```{r}
corrplot(cor(players_num[col2]), type = 'lower',method = 'shade', order = 'hclust')
players_model <- subset(players_norm)
players_model$player_positions <- c(players_22$player_positions)
head(players_model)
```




```{r}
#here we do the cool violin plots to check distributions
p <- ggplot(data = melt(players_num), aes(y = variable, x = value, fill = variable, alpha = 0.7)) + 
             geom_boxplot() + geom_violin() + scale_fill_manual(values = viridis(45)) + guides(fill = "none")
p
```

```{r}
players.pca<-prcomp(players_num,center=TRUE, scale.=TRUE)
summary(players.pca)
players.pca
fviz_eig(players.pca, addlabels = TRUE)
```

```{r}
fviz_pca_var(players.pca, labelsize = 2, alpha.var = 1.0, title = "Factor Plane for the FIFA 22 Data")
```

#CLASSIFICATION











#CLASSIFICATION


### train-test split

```{r}
## 70% of the sample size
smp_size <- floor(0.7 * nrow(players_model))

## set the seed to make your partition reproducible
train_ind <- sample(seq_len(nrow(players_model)), size = smp_size)

train <- players_model[train_ind, ]
test <- players_model[-train_ind, ]
dim(test)
colnames(train)
```

We extract 'club_position' column of train dataset because it will be used as 'cl' argument in knn function.

```{r}


#extract 39th column of train dataset (club_position) that will be used as 'cl' argument in knn function

train_y <- as.factor(train[,39])

#extract 18th column of test set to measure accuracy
test_y <- as.factor(test[,39])
```


``` {r}
##run knn function
test_pred <- knn(train,test,cl=train_y,k=5)
head(test_pred)
```


``` {r}
df_pred=data.frame(test_category,test_pred)
head(df_pred)
```


``` {r}
#create confusion matrix
table <- table(test_category,test_pred)
table
```


``` {r}
#Evaluate the model performance
install.packages("gmodels")
library(gmodels)
CrossTable(x=test_category, y=test_pred,prop.chisq = FALSE)
```


``` {r}

##this function divides the correct predictions by total number of predictions that tell us how accurate the model is.
accuracy <- function(x){sum(diag(x)/(sum(rowSums(x)))) * 100}
accuracy(table)

```


Neural Network



```{r}
library(tidyverse) 
library(neuralnet)

```
Since the club_position column consists of categorical variables, we need to convert the column into a factor

```{r}

players_model <- players_model %>%  
  mutate(club_position=as_factor(club_position))
```

split

```{r}
training_data_rows <- floor(0.70 * nrow(players_model))
set.seed(123) 
training_indices <- sample(c(1:nrow(players_model)), training_data_rows)

```

split d

```{r}
training_data <- players_model[training_indices,] 
test_data <- players_model[-training_indices,]
```



```{r}
print(head(players_model))

```
We build our NN

```{r}
nn=neuralnet(club_position ~ age + height_cm + weight_kg + pace + attacking_crossing + attacking_heading_accuracy + skill_fk_accuracy + skill_long_passing + movement_agility + power_shot_power + power_jumping + power_stamina + power_strength + mentality_aggression + mentality_interceptions + mentality_penalties + mentality_composure
,  
             data=training_data, hidden=c(2,2), linear.output = FALSE)
```

plot nn

```{r}
plot(nn)
```
Testing


```{r}
print(unique(as.character(players_model$club_position)))
```


```{r}

predict <- function(data){ 
  prediction <- data.frame(neuralnet::compute(nn,  

                                              data.frame(data[,-5]))$net.result) 
  labels <- c("RW" , "ST",  "LW" , "RCM" , "CF" , "CDM" , "LCB" , "RDM" , "RS" ,  "LCM" , "SUB", "CAM", "RCB", "LDM", "LB",  "RB",  "LM",  "RM",  "LS",  "CB",  "RES", "RWB", "RF",  "CM",  "LWB", "LAM", "LF",  "RAM" ) 
  prediction_label <- data.frame(max.col(prediction)) %>%  
    mutate(prediction=labels[max.col.prediction.]) %>%  
    select(2) %>%  
    unlist() 

  table(data$club_position, prediction_label) 
}

```

let us check the performance on our training data

```{r}

predict(training_data)

```

EXperiments to reduce classes

[1] "RW"  "ST"  "LW"  "RCM" "CF"  "CDM" "LCB" "RDM" "RS"  "LCM" "SUB" "CAM" "RCB" "LDM" "LB"  "RB"  "LM"  "RM"  "LS"  "CB"  "RES" ""    "RWB"
[24] "RF"  "CM"  "LWB" "LAM" "LF"  "RAM"

```{r}
head(players_22[players_22[,"club_position"] == "RES", "short_name"])
players_22[players_22[,"club_position"] == "", "short_name"]
                
```

```{r}
head(players_full, n=50)
summary(players_full)
```
```{r}
single_positions <- strsplit(players_full$player_positions, split = ",")

typeof(single_positions[1][2])
as.list(single_positions[1])[2]



```
```{r}
install.packages("stringr")     # Install & load stringr package
library("stringr")
goalkeepers <- str_detect(players_22$club_position, "GK")
sum(goalkeepers)

players_22<-subset(players_22, club_position!="GK")
which(apply(X = players_22, MARGIN = 2, FUN = anyNA) == TRUE) # check for NA
dim(players_22)
#head(players_22, n = 2)


```
############################################
SL PROJECT

We start by loading the libraries we will need.

```{r}
library(knitr) 
#library(kableExtra) #nice html tables
#library(magick)     #working with images
library(ggplot2)    #graphs
library(reshape2)   #dataframe reshaping
library(viridis)    #colors
library(stringr)    #working with strings
library(FactoMineR) #factor analysis (PCA)
library(factoextra) #factor analysis (PCA)
#library(kohonen)    #self-organised maps/SOM 

# libraries
library(dplyr)
library(ggplot2)
library(MASS)
#library(randomForest)
library(class)
#library(caret)
install.packages("corrplot")

library(corrplot)
library(gridExtra)
remove.packages(rlang)
install.packages("rlang")
install.packages("https://cran.rstudio.com/bin/windows/contrib/4.1/caret_6.0-92.zip")
install.packages("caret")
library(mlbench)
library(caret)
library(randomForest)

```

We read the file and check the dimensionality of the data

```{r}

players_full <- read.csv("C:\\Users\\gioda\\Downloads\\archive (1)\\players_22.csv") #full dataframe
dim(players_full) #full dataset
```

We use 'apply' function to check for features that lack some data

```{r}
which(apply(X = players_full, MARGIN = 2, FUN = anyNA) == TRUE)
```

Just 16 features and some of them are  not relevant to our purpose.
There is a great lot of various positions in soccer. Here is position classification by FIFA.

```{r}

unique(players_full$club_position)

```
R and L — right and left, F and B — forward and back, C — center, S - striker, M stands for midfielder.

We drop "SUB" and "RES", that stands for "substitutes" and "reserves", and "", because the club position is not specified for these players. Moreover 

```{r}

install.packages("stringr")     # Install & load stringr package
library("stringr")
sub <- str_detect(players_full$club_position, "SUB")
sum(sub)

res <- str_detect(players_full$club_position, "RES")
sum(res)

no_pos <- str_detect(players_full$club_position, "")
sum(no_pos)
```


```{r}


players_pos <-players_full[!(players_full$club_position=="SUB" | 
                      players_full$club_position=="RES" | players_full$club_position=="" | 
                      players_full$club_position=="GK" ),]
dim(players_pos)
dim(players_full[(players_full$club_position=="SUB"|players_full$club_position=="RES"),])
dim(players_full)
```


```{r}
top_leagues <- (players_pos$league_level == 1)
sum(top_leagues, na.rm = TRUE)

```
skills 

```{r}
skills_vars <- c(38:72)
names(players_full[,skills_vars])

```

Top leagues players with defined club position

```{r}

players <- players_pos[top_leagues, skills_vars]
players <- na.omit(players)
dim(players)

```




```{r}

```

